---
title: "Solution Intro-3"
author: "Dr. Osama Mahmoud"
date: "05/07/2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = TRUE)
```

## Exercise 1:
  
  a) consider the following data - similar to the example used for `practical Intro-1` - to generate a data frame named `DT`:
  
```{r, eval=TRUE, echo=FALSE}
DT = data.frame(id = 1:10,
                ht=c(155, 152, 164, 175, 193, 203, 190, 183, 155, 169),
                wt=c(80, 85, 45, 69, 86, 110, 106, 96, 90, 89),
                gender=c("m", "m", "f", "m", "f", "f", "f", "m", "f", "m"))
knitr::kable(DT)
```

```{r,}
DT = data.frame(id = 1:10,
                ht=c(155, 152, 164, 175, 193, 203, 190, 183, 155, 169),
                wt=c(80, 85, 45, 69, 86, 110, 106, 96, 90, 89),
                gender=c("m", "m", "f", "m", "f", "f", "f", "m", "f", "m"))
DT
```
  
  b) Add a new categorical variable `bmi.grp` to the data frame which is defined as follows:

  * bmi $\leq$ 18.5 $\rightarrow$ underweight

  * 18.5 < bmi $\leq$ 25 $\rightarrow$ normal

  * 25 < bmi $\leq$ 30 $\rightarrow$ overweight 

  * 30 < bmi $\rightarrow$ obesity
  
```{r}
DT$bmi = DT$wt/(DT$ht/100)^2
DT$bmi.grp = ifelse(DT$bmi <= 18.5, "underweight", "obesity")
DT$bmi.grp = ifelse(DT$bmi > 18.5 & DT$bmi <= 25, "normal", DT$bmi.grp)
DT$bmi.grp = ifelse(DT$bmi > 25 & DT$bmi <= 30, "overweight", DT$bmi.grp)
```

  c) Generate a vector `z` consisting of the first 30 elements of the Fibonacci series. By definition, the first two elements of this series are 0 and 1. All further elements are calculated as sum of the preceding two elements, so $z=0,1,1,2,3,5,\ldots$

```{r}
# Manually set the first two elements
z = c(0,1)
for(i in 2:29){
  z[i+1] = z[i] + z[i-1]
}
z
```

## Exercise 2: 

  a) load the internal R data set `mtcars` and view its help page to find out abouts its varaible description.

```{r}
data(mtcars)
str(mtcars)
?mtcars
```

  b) Fit a regression model of "Number of car cylinders"" on "Miles per gallon" using the function `lm(mpg ~ cyl, data=mtcars)` and assign the `Model` for that object. Then show `summary` of that object.
  
```{r}
Model = lm(mpg ~ cyl, data=mtcars)
summary(Model)
```

  c)  Write a function to extract the effect estimate, standard error and the p-value from a linear regression model. Call the function `beta_se`. Hint: use the regression model object as the input and extract summary(model)$coefficient.
  
```{r}
# to explore what names can be extracted from the summary object.
names(summary(Model))
# The function
beta_se = function(Simple.Reg.model){
  Summary.model = summary(Simple.Reg.model)
  Out = Summary.model$coefficients[2,c("Estimate", "Std. Error", "Pr(>|t|)")]
  return(Out)
}
# Implement the `beta_se` function
beta_se(Model)
```
  
  d) Use the `beta_se` function to extract the effect estimates, standard errors and p-values of the regression models of (1)"Displacement" on "Miles per gallon" and (2) "Rear axle ratio" on "Miles per gallon".
  
```{r}
Model1 = lm(mpg ~ disp, data = mtcars)
beta_se(Model1)
Model2 = lm(mpg ~ drat, data = mtcars)
beta_se(Model2)
```